---
title: "Homework 2"
author: "Joshua Carpenter"
date: "2023-10-03"
output: github_document
---

```{r setup, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
```


## Problem 1
```{r five38, include=FALSE}
pols <- 
  read_csv("Data/pols-month.csv") %>%
  separate(mon, c("year", "month"), sep="-", extra = "drop") %>%
  mutate(month = month.name[as.numeric(month)], 
         president = case_when(prez_dem == 1 ~ "dem", 
                               prez_gop == 1 ~ "gop")) %>%
  select(-starts_with("prez"))

snp <-
  read_csv("Data/snp.csv") %>%
  separate(date, c("month", NA, "year"), sep="/") %>%
  mutate(year = paste(ifelse(year < 23, "20", "19"), year, sep = ""), 
         month = month.name[as.numeric(month)]) %>%
  select(year, month, close)

unemploy <-
  read_csv("Data/unemployment.csv") %>%
  pivot_longer(Jan:Dec, 
               names_to = "month", 
               values_to = "unemploy",
               values_drop_na = TRUE) %>%
  transmute(year = as.character(Year), 
            month = month.name[match(month, month.abb)], 
            unemploy)

five38 <-
  list(pols, snp, unemploy) %>%
  reduce(full_join, by = join_by(year, month))
```

The below code creates a dataset combining several datasets from a FiveThirtyEight article called "Science Isn’t Broken: It’s just a hell of a lot harder than we give it credit for." The result is a dataset with `r nrow(five38)` observations containing information about the number of national politicians in the US from each party, the closing values of the S&P stock index, and percent unemployment in the US for each month between `r five38[1, "month"]` `r five38[1, "year"]` and `r five38[nrow(five38), "month"]` `r five38[nrow(five38), "year"]`. There are complete data for January 1950 to June 2015 except that August to December 1974 has no president data due to a mistake in the raw data. The variable `president` indicates whether the president was democratic ("dem") or republican ("gop") during that month, `close` indicates the closing price of the S&P stock index on one day that month, and `unemploy` gives the percent unemployment rate.

```{r, eval=FALSE}
<<five38>>
```

Here are the rows of the resulting dataset corresponding to the year 1950:
```{r, echo=FALSE}
five38 %>%
  filter(year == "1950") %>%
  kable()
```

## Problem 2
```{r trash_wheel, include=FALSE}
mr_trash <-
  read_excel("Data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Mr. Trash Wheel", 
             skip = 2,
             col_names = c("dumpster", "month", "year", "date", "weight", 
                           "volume", "bottles_plastic", "polystyrene", 
                           "cigarettes", "bottles_glass", "bags", "wrappers", 
                           "balls", "homes", "drop1", "drop2")) %>%
  select(-starts_with("drop")) %>%
  filter(!is.na(dumpster)) %>%
  mutate(wheel_name = "Mr", 
         year = as.numeric(year), 
         homes_calc = weight * 500 / 30)  %>%
  select(wheel_name, everything())

prof_trash <-
  read_excel("Data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Professor Trash Wheel", 
             skip = 2,
             col_names = c("dumpster", "month", "year", "date", "weight", 
                           "volume", "bottles_plastic", "polystyrene", 
                           "cigarettes", "bottles_glass", "bags", "wrappers", 
                           "homes")) %>%
  filter(!is.na(dumpster)) %>%
  mutate(wheel_name = "Professor", homes_calc = weight * 500 / 30) %>%
  select(wheel_name, everything())

gwynnda <-
  read_excel("Data/202309 Trash Wheel Collection Data.xlsx", 
             sheet = "Gwynnda Trash Wheel", 
             skip = 2,
             col_names = c("dumpster", "month", "year", "date", "weight", 
                           "volume", "bottles_plastic", "polystyrene", 
                           "cigarettes", "bags", "wrappers", 
                           "homes")) %>%
  filter(!is.na(dumpster)) %>%
  mutate(wheel_name = "Gwynnda", homes_calc = weight * 500 / 30) %>%
  select(wheel_name, everything())

trash_wheel <- 
  bind_rows(mr_trash, prof_trash, gwynnda) %>%
  group_by(wheel_name)
```
```{r, include=FALSE}
total_weight_prof <- trash_wheel %>%
  filter(wheel_name == "Professor") %>%
  summarise(total_weight = sum(weight)) %>%
  select(total_weight) %>%
  as.numeric()

cigarettes_gwynnda_202107 <- trash_wheel %>%
  filter(wheel_name == "Gwynnda", year == 2021, month == "July") %>%
  summarise(total_cigs = sum(cigarettes)) %>%
  select(total_cigs) %>%
  as.numeric()
```

The below code code produces a combined dataset containing trash collection data for "Mr. Trash Wheel", "Professor Trash Wheel", and "Gwynnda, the Good Wheel of the West" since each of their construction in 2015, 2017, and 2021 respectively up to June 2023. You can read more about the trash wheel family [here](https://www.mrtrashwheel.com/). The dataset contains `r nrow(trash_wheel)` total observations, grouped by `wheel_name` and ordered by `date`. The other variables indicate the weight and volume of garbage collected each day, as well as the quantity of garbage items of various types. The total weight of trash collected by Professor Trash Wheel was `r round(total_weight_prof)` tons and the total number of cigarettes collected by Gwynnda in July 2021 was `r round(cigarettes_gwynnda_202107) %>% format(scientific = FALSE, big.mark = ",")`.

```{r, eval=FALSE}
<<trash_wheel>>
```