---
title: "Homework 2"
author: "Joshua Carpenter"
date: "2023-10-03"
output: github_document
---

```{r setup, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
```


## Problem 1
```{r five38, include=FALSE}
pols <- 
  read_csv("Data/pols-month.csv") %>%
  separate(mon, c("year", "month"), sep="-", extra = "drop") %>%
  mutate(month = month.name[as.numeric(month)], 
         president = case_when(prez_dem == 1 ~ "dem", 
                               prez_gop == 1 ~ "gop")) %>%
  select(-starts_with("prez"))

snp <-
  read_csv("Data/snp.csv") %>%
  separate(date, c("month", NA, "year"), sep="/") %>%
  mutate(year = paste(ifelse(year < 23, "20", "19"), year, sep = ""), 
         month = month.name[as.numeric(month)]) %>%
  select(year, month, close)

unemploy <-
  read_csv("Data/unemployment.csv") %>%
  pivot_longer(Jan:Dec, 
               names_to = "month", 
               values_to = "unemploy",
               values_drop_na = TRUE) %>%
  transmute(year = as.character(Year), 
            month = month.name[match(month, month.abb)], 
            unemploy)

five38 <-
  list(pols, snp, unemploy) %>%
  reduce(full_join, by = join_by(year, month))
```

In this problem, we create a dataset that combines several datasets from a FiveThirtyEight article called "Science Isn’t Broken: It’s just a hell of a lot harder than we give it credit for." The result is a dataset with `r nrow(five38)` observations containing information about the number of national politicians in the US from each party, the closing values of the S&P stock index, and percent unemployment in the US for each month between `r five38[1, "month"]` `r five38[1, "year"]` and `r five38[nrow(five38), "month"]` `r five38[nrow(five38), "year"]`. There are complete data for January 1950 to June 2015 except that August to December 1974 has no president data due to a mistake in the raw data.

```{r, eval=FALSE}
<<five38>>
```

